# Data

Before analysing any data, it needs to be imported, and converted into the proper structure. In QuaDramA, we process dramatic texts in multiple stages, described below.

## Origin: TEI-XML

The base format that we use (and in which we put all our annotations) is an XML format known as [TEI](https://tei-c.org). This format is used by most researchers doing quantitatitve drama analysis. An excellent source for dramas in the proper format is [DraCor](https://github.com/dracor-org/), maintained by [Frank Fischer](https://www.hse.ru/en/org/persons/182492735).

While we are using GerDraCor as a basis, we have added linguistic annotations to a number of plays, and integrated more plays (e.g., translations) into the corpus. This corpus can be found [here](https://github.com/quadrama/Corpus).

## Preprocessed data

As a first step, we process all dramatic texts using our [DramaNLP](https://github.com/quadrama/DramaNLP) pipeline. The result of this processing is a set of CSV files for each play that contains the information in the play in a format suitable for analysis with R. [This repository](https://github.com/quadrama/data_test) contains two plays in the format, which are also included in our R package.


| File | Description |
| ------------------- | ------------------------------------------------- |
| *ID*.Metadata.csv   | Meta data for the play (author, title, language, ...) |
| *ID*.Characters.csv | Characters of the play, with some character specific information |
| *ID*.Entities.csv   | All discourse entities (including characters, but also all other coreference chains) |
| *ID*.Mentions.csv   | Mentions associated with characters |
| *ID*.StageDirections.csv | The stage directions of the play |
| *ID*.UtterancesWithTokens.csv | All character utterances of the play |
| *ID*.Segments.csv | Information about acts and scenes |
: Different CSV files used in the analysis. *ID* is a placeholder for a unique identifier for the play


## Sample data

```{r}
# Load Emilia Galotti
data(rksp.0)

# Load Miß Sara Sampson
data(rjmw.0)

text <- combine(rksp.0, rjmw.0)

```

## Collection data

To work with corpus data, it's necessary to define a corpus first. In DramaAnalysis, we consider corpora as sets of texts. Typically, these sets have names, such as "comedies", but it does not technically matter why texts are in a collection.

Pre-defined collections can be downloaded with the function `installCollectionData()`. This function clones a git repository ([this one](https://github.com/quadrama/metadata)), which contains a number of plain text files that in turn contain drama ids.

**TODO: ids, corpus prefixes**

Once a vector with ids has been defined, it can be passed as an argument to the function `loadDrama()`. The returning `QDDrama` object contains all loadable plays.

# Basics

This is not really an R tutorial. If you're looking for a full-fledged R tutorial, we recommend [@Arnold:2015aa], because it is very practical, and also deals with text analysis. However, this chapter introduces some concepts related to R.

## R Basics

### Variable assignment

Assigning variables is one of the things you do most frequently in most programming languages. In R, this is done with the arrow operator: `a <- 5`. This creates a new variable `a` that has the value 5. Conceptually, variables are aliases for values. The same value can be assigned to multiple names, but one variable only can point to one value. 

### Function calls

The R package for drama analysis defines a number of functions. Functions are mini programs that can be executed anytime by a user. Function calls (i.e., the execution of a function) can be recognized by round parentheses: `sessionInfo()` is a function call. If you enter this in the R console and press enter, the function gets executed and its return value is printed on the console. The function [`sessionInfo()`](https://www.rdocumentation.org/packages/utils/versions/3.6.0/topics/sessionInfo) can be used to get information about your R session and installation. The function does not take any arguments, which is why the round parentheses are empty. The function `sum`, on the other hand, takes arguments: The values it should add. Thus, calling `sum(1,1)` prints 2 on the console. 


## magrittr Pipes / `%>%` {#magrittr}

In the tutorial, we make heavy use of `magrittr`-pipes. These are provided by the R package [`magrittr`](https://cran.r-project.org/web/packages/magrittr/vignettes/magrittr.html). The core idea behind pipes is to use the output of one function directly as input for the next function -- to create a pipeline of functions (if you're familiar with the unix command line you probably have used this before).

The pipeline is represented by `%>%`. The semantics of `%>%` is to use the output of whatever comes before as *first argument* for whatever comes after. Thus, if we write `runif(5) %>% barplot`, the result of `runif(5)` (which is a vector with five random numbers) is used as the first argument for the function `barplot()`. This is eqivalent to:

```{r eval=FALSE}
r <- runif(5)
barplot(r)
```

The nice thing about magrittr is that you can create pipes that are much longer:

```{r eval=FALSE}
d <- loadDrama("test:rksp.0")

d %>%
  dictionaryStatistics() %>%
  filter(d) %>%
  format(d) %>%
  barplot() 
```

Again, this is equivalent to a version with variables:

```{r eval=FALSE}
d <- loadDrama("test:rksp.0")

ds <- dictionaryStatistics(d)
dsf <- filter(ds, d)
dsff <- format(dsf, d)
barplot(dsff)
```

Pipes are much faster to write and more transparent, that's why they are used in this tutorial.
