# Character Exchange

Character exchange -- entering and leaving the stage -- often takes place at scene boundaries. We offer multiple ways to measure and visualize this exchange over the course of the play.

## Hamming Distance

The [hamming distance](https://en.wikipedia.org/wiki/Hamming_distance) has been introduced in information theory by Richard Hamming [@Hamming:1950aa]. Intuitively, the hamming distance expresses the number of characters that either enter or leave the stage. Its application to a dramatic text is straightforward, using the function `hamming()`. By default, the hamming distance is calculated over scene boundaries.

```{r}

hamming(rksp.0, variant="Hamming")

```

The return value of the function is just a vector of numbers, one less than the number of scenes in the play. Of course, this can be plotted:

```{r}
plot(hamming(rksp.0, variant="Hamming"),
     type="l",
     ylab="Hamming distance",
     xlab="Scenes")
```


Hamming distance, in its original definition as well as the `Hamming` variant in our function, returns absolute values. As usual, these values should be normalized, and the function `hamming()` provides two ways of doing that.


### NormalizedHamming

Straightforwardly, one can normalise by the overall number of characters in the play. This value is only maximal, if at once scene boundary, one subset of characters enters the stage, and the entire other subset leaves the stage. As we can see below, this is not the case in our demo text.

```{r}

hamming(rksp.0, variant="NormalizedHamming")

```

### Trilcke

One more possible normalization has been proposed by @Trilcke:2017ab. Instead of normalizing with all characters in the play, the Trilcke variant only normalizes with the characters in the two adjacent scenes. Thus, if all characters leave the stage, and a new set of characters enter it, the distance is maximal. This however does not have to include *all* characters in the play.

```{r}

hamming(rksp.0, variant="Trilcke")

```

The Trilcke variant is the default setting of the `hamming()`-function.


## Corpus Analysis



```{r warning=FALSE}
sturm_und_drang.ids <- c("qd:11f81.0", "qd:11g1d.0", "qd:11g9w.0", 
                         "qd:11hdv.0", "qd:nds0.0",  "qd:r12k.0", 
                         "qd:r12v.0",  "qd:r134.0",  "qd:r13g.0", 
                         "qd:rfxf.0",  "qd:rhtz.0",  "qd:rhzq.0", 
                         "qd:rj22.0",  "qd:tx4z.0",  "qd:tz39.0", 
                         "qd:tzgk.0",  "qd:v0fv.0",  "qd:wznj.0", 
                         "qd:tx4z.0",  "qd:rfxf.0")

sturm_und_drang.plays <- loadDrama(sturm_und_drang.ids)

```

Since determining character exchanges at scene boundaries is not a multi-play task, the function `hamming()` does not allow `QDDrama` objects that contain multiple plays. It's still possible to calculate the exchanges for many plays though, using `split()` and `lapply()`. `split()` is a function defined in the `DramaAnalysis` package and can be used to split up a `QDDrama` object if it contains multiple plays. It returns a list of individual objects. This list is then fed into the function `lapply()`, which applies the function `hamming()` on each element of the list.

```{r}
profiles <- lapply(split(sturm_und_drang.plays), hamming, variant = "Trilcke")
```

The result of this is a list that contains an entry for each play, which in turn reflects the distances at specific scene boundaries. Thus, the lists have different lengths, as the plays have different lengths. We apply the same trick with `lappy()` to create plots of all the plays. Instead of calling a pre-existing function (as before with `hamming()`), we define a new (anonymous) function. This is mainly to allow printing the authors and names of the play into the plots.

```{r fig.height=25, results='hide'}
par(mfrow=c(length(profiles),1), mar=c(2,2,2,2))

lapply(names(profiles), 
       function(x) {
         plot(profiles[[x]],  # take the data
              type="l",       # we print lines
              ylab="Character exchange rate", # y axis label 
              ylim=c(0,1),                    # y axis scale
              # create nice labels
              main=paste(sturm_und_drang.plays$meta[drama==x,]$Name,
                         sturm_und_drang.plays$meta[drama==x,]$documentTitle, 
                         sep=": ", collapse=": ")
         )
       }
)

```
